---
title: "p8105_hw1_bl114"
author: Ben Lebwohl
date: July 2021
output: github_document
---


# Homework 1

### Problem 1

Create a data frame comprised of:

  * a random sample of size 10 from a standard Normal distribution
  * a logical vector indicating whether elements of the sample are greater than 0
  * a character vector of length 10
  * a factor vector of length 10, with 3 different factor “levels”

```{r dataframe}
library(tidyverse)
dataframe_problem_1 = tibble (
  random_sample = rnorm(10),
  logical_vector = random_sample >0,
  charac_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  factor_vector = factor(c("S", "M", "L", "S", "M","L","S", "M", "L", "S"))
)

dataframe_problem_1

```

Try to take the mean of each variable in your dataframe. What works and what doesn’t?

```{r means}

mean(dataframe_problem_1$random_sample)
mean(dataframe_problem_1$logical_vector)
mean(dataframe_problem_1$charac_vector)
mean(dataframe_problem_1$factor_vector)

```


Running the mean function on the random sample and the logical vector works, but running it on the character vector and factor vector does not.

Now, let's apply as.numeric function to the logical, character, and factor variables. We will show the code but not the output:

```{r as.numeric, eval=FALSE}

as.numeric(dataframe_problem_1$logical_vector)
as.numeric(dataframe_problem_1$charac_vector)
as.numeric(dataframe_problem_1$factor_vector)
```

Logical vectors are converted to 0's and 1's (though they are already treated as such when applying the mean function). Character vectors cannot be converted, and factor vectors are converted to 1, 2, and 3 (perhaps alphabetically).


In a second code chunk, we will:

  * Convert the logical vector to numeric, and multiply the random sample by the result
  * Convert the logical vector to a factor, and multiply the random sample by the result
  * Convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r convertions}

logical_numeric = as.numeric(dataframe_problem_1$logical_vector)
dataframe_problem_1$random_sample*logical_numeric

logical_factor = as.factor(dataframe_problem_1$logical_vector)
dataframe_problem_1$random_sample*logical_factor

factor_now_numeric = as.numeric(logical_factor)
dataframe_problem_1$random_sample*factor_now_numeric

```

### Problem 2

Load the dataset "penguins"

```{r penguins}
data("penguins", package = "palmerpenguins")
```
The penguins data set has the following variables: `r colnames(penguins)`

The penguins dataset contains `r nrow(penguins)` rows

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`


Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable. Export your first scatterplot to your project directory using ggsave.

```{r scatterplot}
ggplot(penguins, aes (x=bill_length_mm, y=flipper_length_mm, color = species)) + geom_point()
ggsave("species_scatterplot.pdf")

```



